\pkg{Rgbp}: An R Package for Hierarchical Modeling and Method Checking

### 2013-07-17 jss/article.tex
proofreading

%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an R package for estimating and validating a two-level model (a random-effects mixed model). The estimation procedure utilizes Bayesian machinery and the validation involves checking frequency properties of the procedure via repeated sampling (which we call ``method checking"). It is found that \pkg{Rgb} characterized by such Bayesian-frequentist reconciliation has strong points of both sides, working well in small samples and yielding good coverage probability. Following three examples for Gaussian, Poisson and Binomial data will validate this point.}
\Keywords{hierarchical model, multilevel model, random effects mixed model, method checking, coverage probability, Normal, Poisson, Binomial,  shrinkage, \proglang{R}}
\Plainkeywords{keywords, comma-separated, not capitalized, r} %% without formatting
%% at least one keyword must be supplied

@@

%\subsection{Normal: 8 Schools} \label{8schooldata}
%The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \cite{1981}.
%\begin{CodeChunk}
%\begin{CodeInput}
%R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
%R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
%\end{CodeInput}
%\end{CodeChunk}

@@

%\subsection{Poisson: 31 Hospitals}
%The following data are from medical profiling evaluations for Coronary Artery Bypass Graft (CABG) surgeries of 31 New York hospitals conducted in 1992 \cite{2012}. It comprises of the number of deaths within a month of CABG surgeries in each hospital ($z_{j},~j=1, \ldots, 31$) and the total number of patients receiving CABG surgeries (case load) in each hospital ($n_{j},~j=1, \ldots, 31$). The following code is an example of input based on the last ten hospital data.
%\begin{CodeChunk}
%\begin{CodeInput}
%R> z <- c( 14,   9,  15,  13,  35,  26,  25,  20,   35,   27)
%R> n <- c(593, 602, 629, 636, 729, 849, 914, 940, 1193, 1340)
%\end{CodeInput}
%\end{CodeChunk}
%Considering the type of data, it is reasonable to assume the number of deaths in each hospital follow independent Poisson distributions given an unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where, for each $j$, $n_{j}$ can be interpreted as an exposure (not necessarily an integer) and $\lambda_{j}$ as the true death rate per exposure. This data set is also included in the package and can be called by `\code{data(hospital)}' on \proglang{R}.

%\subsection{Binomial: 18 Baseball Hitters}
%The following dataset contains information on the batting averages of 18 major league baseball players through their first 45 official at-bats of the 1970 season \cite{1975}. In addition one covariate relating to the position each player was playing was observed and for illustrative purposes, we transform this variable into a binary indicator (1 if a player was an outfielder and 0 otherwise).
%\begin{CodeChunk}
%\begin{CodeInput}
%R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  %8,  7)
%R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, %45, 45)
%R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
%\end{CodeInput}
%\end{CodeChunk}
%The data indicate that independent Binomial distribution is appropriate for each player's number of hits among 45 at-bats conditioning on the unknown true batting average $p_{j}$: $z_{j}\vert p_{j}\stackrel{ind}{\sim} \textrm{Binomial}(n_{j}, p_{j}), ~j=1, \ldots, 18$. This data set is also a part of the package and can be called on \proglang{R} by `\code{data(baseball)}'.

\section[Multilevel Structure]{Multilevel Structure} \label{multilevel}
A two-level or multilevel model, also called a conditionally independent hierarchical model \cite{1989}, is a very powerful tool for exploring the hierarchical structure in data. For example, we can imagine that there exists a district-level hierarchy (bigger population) for observed school-level data, or a state-level hierarchy for observed hospital-level data in a certain state.

@@

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
In our examples the hyper-prior distribution is an assumed distribution on the second-level parameters and with the goal of objectivity in mind \code{gbp} assumes non-informative hyper-prior distributions.
\begin{equation}
\beta \sim \textrm{Uniform on}~ \mathbf{R}^{m}~~\textrm{and}~~A \sim \textrm{Uniform}(0, \infty) ~~(\textrm{or} ~\frac{1}{r}\sim \textrm{Uniform}(0, \infty)),
\end{equation}

@@

\subsection[Shrinkage Estimation]{Shrinkage Estimation}
Estimating the shrinkage factors ($B_1, \ldots B_k$) is the key estimation problem with the hierarchical models \code{gbp} assumes. As we can see in (\ref{normalpost}), (\ref{gammapost}), and (\ref{betapost}), the posterior means are a linear function of the shrinkage factors and the posterior variances are also linear (Gaussian), quadratic (Poisson), or cubic (Binomial) functions. A natural method then to estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$ is to first estimate the shrinkage factors.
%TODO: If we cite somgthing, (Morris, 2011) might be appropriate format to me, but in our case it looks Morris (2011), which is confusing.

\subsection[ADM]{Adjustment for Density Maximization}
It is noted that the shrinkage factors ($B_1, \ldots B_k$) are a function of the second-level variance component, \emph{i.e.}, $B_{j}\equiv V_{j}/(V_{j}+A)=B_{j}(A)$ for Gaussian and $B_{j}\equiv r/(r+n_{j})=B_{j}(r)$ for Poisson and Binomial models. Current popular methods of estimation via maximization (MLE or MAP) of these unknown shrinkage factors can result in estimates lying on the boundary of the parameter space. In the Normal-Normal model this situation typically arises when the within-group variation is greater than the between-group variation resulting in a parameter estimate of $\hat{A} = 0$ \cite{2012}.

To continue with a maximization-based estimation procedure but to steer clear of aforementioned issues we make use of adjustment for density maximization (ADM) \cite{2011}. In general ADM is a procedure to obtain estimates of posterior moments via maximization if the underlying posterior can be approximated by a Pearson family. For our purposes we can approximate the posterior distribution of a shrinkage factor with the Beta distribution allowing us to finally obtain estimates of the posterior moments, \emph{i.e.}, $E(B_{j}\vert\textrm{data})$ and $Var(B_{j}\vert\textrm{data})$, without any trouble that MLE can cause. Please refer to \cite{2011} for additional advantages of ADM.

@@

\subsection[Approximation to Posterior Distribution by Moment Matching]{Approximation to Posterior Distribution via Matching Moments}
After estimating the two posterior moments, for example $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$, \code{gbp} reasonably approximates a posterior distribution of the mean effects given the data by assuming a reasonable distribution and matching moments. For the Binomial-Beta model we approximate $p_{j}\vert \textbf{z}$ with a Beta distribution and for the Poisson-Gamma model we approximate $\lambda_{j}\vert \textbf{z}$ with a Gamma distribution. Finally for the Normal-Normal model we actually estimate the first three moments and approximate $\mu_{j}\vert \textbf{y}$ with a skewed-t distribution.

@@

% TODO: Here, Christiansen and Morris (1997) seems appropriate format
There are two kinds of model justification processes; one is a model checking and the other is a method checking. The model checking is for the justification of the multilevel model on a specific dataset. One possible question is, ``Can this dataset benefit from such a multilevel modeling?'' \cite{1997} answered this question by using the Negative-Binomial mixture model on Poisson data to justify the second-level hierarchy. They found that their data had more variation than expected of the first-level Poisson distribution and Poisson hierarchical model could successfully account for such additional variation.

@@

For reference, the explanation will be based on Normal-Normal model because the idea can be easily applied to the other two models.

@@

\subsubsection{Rao-Blackwellized Unbiased estimates of Coverage Probabilities}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking conditional expectation given a sufficient statistic. Based on this idea, we define $E(I_{ij}\vert y_{ij}, A, \beta$), where $A$ and $\beta$ were given at first (see 4.1) and $y_{ij}$ is a sufficient statistic. This expectation is the same as Pr$(\hat{\mu}_{ij, low}\le \mu_{ij} \le\hat{\mu}_{ij, upp}\vert y_{ij}, A, \beta)$, where ($\hat{\mu}_{ij, low}$, $\hat{\mu}_{ij, upp}$) is the $j$-th group's interval estimate on the $i$-th dataset. We can calculate this probability  because we know the distribution of $\mu_{ij} \vert y_{ij}, A, \beta$ in (\ref{normalpost}). Note that conditioning on $y_{ij}$ is equivalent to conditioning on $\mathbf{y}$ as long as $A$ and $\beta$ are known. Then, we can estimate the first group's coverage probability by $\sum_{i=1}^{500}E(I_{i1}\vert y_{i1}, A, \beta)/500$, which is also unbiased but more accurate than the previous simple estimator.

@@

Assuming an additional hierarchy is reasonable, a model-fitting process begins.  We know the true death rate per exposure after CABG surgery might be small and the caseload ($n_{j}$) is much bigger than the number of deaths ($z_{j}$) in each hospital. Then the independent Poisson distribution, \emph{i.e.},  $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ would be the first choice to describe the uncertainty in our data. Note that caseload  ($n_{j}$) can be interpreted as an exposure, which is not necessarily an integer.

@@

Also, \code{gbp} can give us interval estimates with different confidence level, for example 90\%,  and the appropriate code adjustment is \code{R> p <- gbp(z, n, Alpha = 0.9, mean.PriorDist = 0.02, model = "pr")}.  Then, the function \code{coverage} will evaluate whether interval estimates achieve 90\% confidence level.

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
% TODO: As stated in ??
As stated in \ref{8schooldata} the Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \cite{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}

@@

The following dataset from the New York Times published on 26 April 1970 contains information on the batting averages of 18 major league baseball players through their first 45 official at-bats of the 1970 season \cite{1975}. In addition one covariate relating to the position of each player was observed and for illustrative purposes, we transform this variable into a binary indicator (1 if a player was an outfielder and 0 otherwise).
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  8,  7)
R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45)
R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
\end{CodeInput}
\end{CodeChunk}

@@

         obs.mean  n   X1 prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.400 45 1.00      0.310     0.715    0.248     0.335    0.429   0.046
2           0.378 45 1.00      0.310     0.715    0.244     0.329    0.420   0.045
3           0.356 45 1.00      0.310     0.715    0.240     0.323    0.411   0.044
4           0.333 45 1.00      0.310     0.715    0.236     0.316    0.403   0.043
5           0.311 45 1.00      0.310     0.715    0.230     0.310    0.396   0.042
6           0.311 45 0.00      0.233     0.715    0.179     0.256    0.341   0.041
7           0.289 45 0.00      0.233     0.715    0.175     0.249    0.331   0.040
8           0.267 45 0.00      0.233     0.715    0.171     0.243    0.323   0.039
9           0.244 45 0.00      0.233     0.715    0.166     0.237    0.315   0.038
10          0.244 45 1.00      0.310     0.715    0.210     0.291    0.379   0.043
11          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
12          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
13          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
14          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
15          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
16          0.200 45 0.00      0.233     0.715    0.155     0.224    0.302   0.038
17          0.178 45 0.00      0.233     0.715    0.148     0.218    0.297   0.038
18          0.156 45 0.00      0.233     0.715    0.140     0.211    0.292   0.039
colMeans    0.265 45 0.44      0.267     0.715    0.191     0.267    0.351   0.041
\end{CodeOutput}
\end{CodeChunk}
Our model reflects on the additional information, \emph{i.e.}, indicator covariate (1 for outfielder and 0 for other positions), estimating two different prior means, 0.310 and 0.233. Also note that shrinkage estimates are the same for all players. It makes sense because shrinkage ($B_{j}\equiv r / (r+n_{j})$) is determined by the relative amount of information between the first-level and the second-level hierarchies. The fact that all players have the same amount of observed information ($n_{j}=45$) and $\hat{r}=\textrm{exp}(4.727)=113$ from below summary tell why the estimated shrinkage is 0.715 (=113 / (113 + 45)).

@@

The 95\% interval plot shows us range of true batting average for each player, which clarifies the regression toward the mean (RTTM) within two groups. Let's see the 10th, 14th, and 15th players on the graph for example. They are outfielders but their observed batting averages are far lower than the first five outfielders. RTTM means that it can happen by their bad luck though in the long run their batting averages will come back to their expected ones as outfielders (blue horizontal line). The fact that their sample means are closer to the lower part of their 95\% intervals supports this argument. So, their posterior means (red dots) can also be interpreted as their RTTM.

@@

In short the \pkg{Rgbp} is an R package that utilizes Bayesian machinery to provide a method of estimating two-level hierarchical models from many families in a fast and computationally efficient manner.

>>>>

%% an abstract and keywords
\Abstract{
\pkg{Rgbp} is an R package that utilizes Bayesian machinery to provide a method of estimating two-level hierarchical models (also called random-effects mixed models) for Normal, Poisson, and Binomial data in a fast and computationally efficient manner. Good frequency property is another feature of \pkg{Rgbp}, which can be validated via its repeated sampling procedure called ``method checking."  It is found that such Bayesian-frequentist reconciliation allows \pkg{Rgbp} to have strong points of both sides, working well in small samples and yielding good coverage probabilities for its interval estimates.}
\Keywords{hierarchical model, multilevel model, random effects mixed model, method checking, coverage probability, Normal, Poisson, Binomial,  shrinkage, \proglang{R}}
\Plainkeywords{keywords, comma-separated, not capitalized, r} %% without formatting
%% at least one keyword must be supplied

@@

%\subsection{Normal: 8 Schools} \label{8schooldata}
%The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \citep{1981}.
%\begin{CodeChunk}
%\begin{CodeInput}
%R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
%R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
%\end{CodeInput}
%\end{CodeChunk}

@@

%\subsection{Poisson: 31 Hospitals}
%The following data are from medical profiling evaluations for Coronary Artery Bypass Graft (CABG) surgeries of 31 New York hospitals conducted in 1992 \citep{2012}. It comprises of the number of deaths within a month of CABG surgeries in each hospital ($z_{j},~j=1, \ldots, 31$) and the total number of patients receiving CABG surgeries (case load) in each hospital ($n_{j},~j=1, \ldots, 31$). The following code is an example of input based on the last ten hospital data.
%\begin{CodeChunk}
%\begin{CodeInput}
%R> z <- c( 14,   9,  15,  13,  35,  26,  25,  20,   35,   27)
%R> n <- c(593, 602, 629, 636, 729, 849, 914, 940, 1193, 1340)
%\end{CodeInput}
%\end{CodeChunk}
%Considering the type of data, it is reasonable to assume the number of deaths in each hospital follow independent Poisson distributions given an unknown true rate parameter $\lambda_{j}$: $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ where, for each $j$, $n_{j}$ can be interpreted as an exposure (not necessarily an integer) and $\lambda_{j}$ as the true death rate per exposure. This data set is also included in the package and can be called by `\code{data(hospital)}' on \proglang{R}.

%\subsection{Binomial: 18 Baseball Hitters}
%The following dataset contains information on the batting averages of 18 major league baseball players through their first 45 official at-bats of the 1970 season \citep{1975}. In addition one covariate relating to the position each player was playing was observed and for illustrative purposes, we transform this variable into a binary indicator (1 if a player was an outfielder and 0 otherwise).
%\begin{CodeChunk}
%\begin{CodeInput}
%R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  %8,  7)
%R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, %45, 45)
%R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
%\end{CodeInput}
%\end{CodeChunk}
%The data indicate that independent Binomial distribution is appropriate for each player's number of hits among 45 at-bats conditioning on the unknown true batting average $p_{j}$: $z_{j}\vert p_{j}\stackrel{ind}{\sim} \textrm{Binomial}(n_{j}, p_{j}), ~j=1, \ldots, 18$. This data set is also a part of the package and can be called on \proglang{R} by `\code{data(baseball)}'.

\section[Multilevel Structure]{Multilevel Structure} \label{multilevel}
A two-level or multilevel model, also called a conditionally independent hierarchical model \citep{1989}, is a very powerful tool for exploring the hierarchical structure in data. For example, we can imagine that there exists a district-level hierarchy (bigger population) for observed school-level data, or a state-level hierarchy for observed hospital-level data in a certain state.

@@

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
The hyper-prior distribution is an assumed distribution on the second-level parameters and with the goal of objectivity in mind \code{gbp} assumes non-informative hyper-prior distributions.
\begin{equation}
\beta \sim \textrm{Uniform on}~ \mathbf{R}^{m}~~\textrm{and}~~A \sim \textrm{Uniform}(0, \infty) ~~(\textrm{or} ~\frac{1}{r}\sim \textrm{Uniform}(0, \infty)),
\end{equation}

@@

\subsection[Shrinkage Estimation]{Shrinkage Estimation}
Estimating the shrinkage factors ($B_1, \ldots, B_k$) is the key estimation problem with the hierarchical models \code{gbp} assumes. As we can see in (\ref{normalpost}), (\ref{gammapost}), and (\ref{betapost}), the posterior means are a linear function of the shrinkage factors and the posterior variances are also linear (Gaussian), quadratic (Poisson), or cubic (Binomial) functions. A natural method then to estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$ is to first estimate the shrinkage factors.

\subsection[ADM]{Adjustment for Density Maximization}
It is noted that the shrinkage factors ($B_1, \ldots B_k$) are a function of the second-level variance component, \emph{i.e.}, $B_{j}\equiv V_{j}/(V_{j}+A)=B_{j}(A)$ for Gaussian and $B_{j}\equiv r/(r+n_{j})=B_{j}(r)$ for Poisson and Binomial models. Current popular methods of estimation via maximization (MLE or MAP) of these unknown shrinkage factors can result in estimates lying on the boundary of the parameter space. In the Normal-Normal model this situation typically arises when the within-group variation is greater than the between-group variation resulting in a parameter estimate of $\hat{A} = 0$ \citep{2012}.

To continue with a maximization-based estimation procedure but to steer clear of aforementioned issues we make use of adjustment for density maximization (ADM) \citep{2011}. In general ADM is a procedure to obtain estimates of posterior moments via maximization if the underlying posterior can be approximated by a Pearson family. For our purposes we can approximate the posterior distribution of a shrinkage factor with the Beta distribution allowing us to finally obtain estimates of the posterior moments, \emph{i.e.}, $E(B_{j}\vert\textrm{data})$ and $Var(B_{j}\vert\textrm{data})$, without any trouble that MLE can cause. Please refer to \cite{2011} for additional advantages of ADM.

@@

\subsection[Approximation to Posterior Distribution by Moment Matching]{Approximation to Posterior Distribution via Matching Moments}
After estimating the two posterior moments, for example $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$, \code{gbp} approximates a posterior distribution of the mean effects given the data by assuming a reasonable distribution and matching moments. For the Binomial-Beta model we approximate $p_{j}\vert \textbf{z}$ with a Beta distribution and for the Poisson-Gamma model we approximate $\lambda_{j}\vert \textbf{z}$ with a Gamma distribution. Finally for the Normal-Normal model we actually estimate the first three moments and approximate $\mu_{j}\vert \textbf{y}$ with a skewed-t distribution.

@@

There are two kinds of model justification processes; one is a model checking and the other is a method checking. The model checking is for the justification of the multilevel model on a specific dataset. One possible question is, ``Can this dataset benefit from such a multilevel modeling?'' \cite{1997} answered this question by using the Negative-Binomial mixture model on Poisson data to justify the second-level hierarchy. They found that their data had more variation than expected of the first-level Poisson distribution and Poisson hierarchical model could successfully account for such additional variation.

@@

From now on, the explanation will be based on Normal-Normal model because the idea can be easily applied to the other two models.

@@

\subsubsection{Rao-Blackwellized Unbiased estimates of Coverage Probabilities}
Rao-Blackwellization improves accuracy of an unbiased estimator by taking conditional expectation given a sufficient statistic. Based on this idea, we define $E(I_{ij}\vert y_{ij}, A, \beta$), where $A$ and $\beta$ were given at first (see 4.1) and $y_{ij}$ is a sufficient statistic. This expectation is the same as Pr$(\hat{\mu}_{ij, low}\le \mu_{ij} \le\hat{\mu}_{ij, upp}\vert y_{ij}, A, \beta)$, where ($\hat{\mu}_{ij, low}$, $\hat{\mu}_{ij, upp}$) is the $j$-th group's interval estimate on the $i$-th dataset. We can calculate this probability  because we know the distribution of ($\mu_{ij} \vert y_{ij}, A, \beta$) in (\ref{normalpost}). Note that conditioning on $y_{ij}$ is equivalent to conditioning on $\mathbf{y}$ as long as $A$ and $\beta$ are known. Then, we can estimate the first group's coverage probability by $\sum_{i=1}^{500}E(I_{i1}\vert y_{i1}, A, \beta)/500$, which is also unbiased but more accurate than the previous simple estimator.

@@

Assuming an additional hierarchy is reasonable, a model-fitting process begins.  We know the true death rate per exposure after CABG surgery might be small and the caseload ($n_{j}$) is much bigger than the number of deaths ($z_{j}$) in each hospital. Then the independent Poisson distribution, \emph{i.e.},  $z_{j}\vert \lambda_{j}\stackrel{ind}{\sim} \textrm{Poisson}(n_{j}\lambda_{j})$, $j=1, \ldots, 31,$ would be the first choice to describe the uncertainty in these data. Note that caseload  ($n_{j}$) can be interpreted as an exposure, which is not necessarily an integer.

@@

Also, \code{gbp} can give us interval estimates with different confidence level, for example 90\%,  and the appropriate code adjustment is \code{R> p <- gbp(z, n, Alpha = 0.9, mean.PriorDist = 0.02, model = "pr")}.  Then, the below code will evaluate whether interval estimates achieve 90\% confidence level.

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}

The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \citep{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}

@@

The following dataset from the New York Times published on 26 April 1970 contains information on the batting averages of 18 major league baseball players through their first 45 official at-bats of the 1970 season \citep{1975}. In addition one covariate relating to the position of each player was observed and for illustrative purposes, we transform this variable into a binary indicator (1 if a player was an outfielder and 0 otherwise).
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10,  9,  8,  7)
R> n <- c(45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45, 45)
R> x <- c( 1,  1,  1,  1,  1,  0,  0,  0,  0,  1,  0,  0,  0,  1,  1,  0,  0,  0) 
\end{CodeInput}
\end{CodeChunk}

@@

         obs.mean  n   X1 prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.400 45 1.00      0.310     0.715    0.248     0.335    0.429   0.046
2           0.378 45 1.00      0.310     0.715    0.244     0.329    0.420   0.045
3           0.356 45 1.00      0.310     0.715    0.240     0.323    0.411   0.044
4           0.333 45 1.00      0.310     0.715    0.236     0.316    0.403   0.043
5           0.311 45 1.00      0.310     0.715    0.230     0.310    0.396   0.042
6           0.311 45 0.00      0.233     0.715    0.179     0.256    0.341   0.041
7           0.289 45 0.00      0.233     0.715    0.175     0.249    0.331   0.040
8           0.267 45 0.00      0.233     0.715    0.171     0.243    0.323   0.039
9           0.244 45 0.00      0.233     0.715    0.166     0.237    0.315   0.038
10          0.244 45 1.00      0.310     0.715    0.210     0.291    0.379   0.043
11          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
12          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
13          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
14          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
15          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
16          0.200 45 0.00      0.233     0.715    0.155     0.224    0.302   0.038
17          0.178 45 0.00      0.233     0.715    0.148     0.218    0.297   0.038
18          0.156 45 0.00      0.233     0.715    0.140     0.211    0.292   0.039
colMeans    0.265 45 0.44      0.267     0.715    0.191     0.267    0.351   0.041
\end{CodeOutput}
\end{CodeChunk}
Our model reflects on the additional information, \emph{i.e.}, indicator covariate (1 for outfielder and 0 for other positions), estimating two different prior means, 0.310 and 0.233. Also note that shrinkage estimates are the same for all players. It makes sense because shrinkage ($B_{j}\equiv r / (r+n_{j})$) is determined solely by the relative amount of information between the first-level and the second-level hierarchies. The fact that all players have the same amount of observed information ($n_{j}=45$) and $\hat{r}=\textrm{exp}(4.727)=113$ from below summary tell why the estimated shrinkage is 0.715 (=113 / (113 + 45)).

@@

The 95\% interval plot shows us range of true batting average for each player, which clarifies the regression toward the mean (RTTM) within two groups. Let's see the 10th, 14th, and 15th players on the graph for example. They are outfielders but their observed batting averages are far lower than the first five outfielders. RTTM means that it can happen by their bad luck though in the long run their batting averages will come back to their expected ones as outfielders (blue horizontal line). So, their posterior means (red dots) can be interpreted as their RTTM.

@@
